---
title: "git_handshake"
output: html_document
---

In this markdown file we will:
1. create a github token (a temporary secret) which allows RStudio interact with your github account.
2. connect that newly acquired token to your Rstudio
3. check if the token works

## 1. First let's a create a token.

- This will open a github website on a 'new token' form
- Give your token a name, like "Natasha Rstudio Github"
- press big green button on the bottom "Generate Token", you will be allowed to copy-paste that token.
- Copy paste the token (it is a string of about 30 characters) and come back here.
- DO NOT close that tab yet, in case you need to copy-paste again

```{r}
# install.packages("usethis") 
# you likely already have 'usethis' library installed. But you can un-comment the line above and run again, if you see an error like:
# Error in library(usethis) : there is no package called ‘usethis’ 

library(usethis)

usethis::git_default_branch_configure(name = "main") # this will do some cleabup first

usethis::create_github_token() # this will open a website in a browser tab.
```

## 2. Allow RStudio to use the token

- We will now share the Token with RStudio, so it can store it in a place where it keeps secrets
- For that we will keep using the usethis library.
- Look in the 'Console' tab on the bottom left. You will be asked a question there.
- if you already had a token, you have AN OPTION to delete and replace it. To do that pick option 2 ('Replace')
- if you never had a token, you will just be asked to paste the long string you copied from github page above. It will look like this "ghp_oxjeoiduh23i8ry3498fh2Q"

```{r}
gitcreds::gitcreds_set() # this (in console) will ask you to Paste the token you Copied from github.
```

## 3. Let's test if it worked

- Now we will check if token got properly saved and if you have 'Write' access to github.
- This means you can now read & write to your public & private repositories.
- you will know if it worked if you see <gitcreds>.... but if you see an Error, something went wrong.

```{r}
gitcreds::gitcreds_get(use_cache = FALSE) # this will either complain, or show your credentials.
```

## 4. What's next? How to make changes to an online github repository:

If you did not see an error above it means you are connected to your github account on this installation of RStudio (this laptop or this server). When you will be in an RProject which is also a code repository, you will see a new tab called 'Git' next to the 'Environment' on top right, with various Git options.

In short:
- change some file in your project, that file will show up in git tab. Only files which got changed or added show up here.
- tick a box next to the file you changed, and select 'commit'. You will be asked to write a short "commit message" which explains what you did. e.g. "fixed a typo in my name". Click "Commit". These changes are still only on your computer.
- now you are ready to send those changes to the ONLINE github repository on the internet. To do that select "Push" with a green arrow up. You can now close any popups.
- your changes should be online, you can go to your repository's website, and you should see them there.

